When and Where: Understanding the Patterns of Emergency Service Calls in Toronto

### 2024-09-27 paper/paper.qmd
Changed the paper structure
Moved acknowledgements to appendix.
Added literature to sociologically discuss the approaches.

The sections are organized as follows: 
@sec-data provides a detailed description of the data and its sources, including the context and ethical considerations. @sec-results presents the results using visualizations as well as discusses the implications of the findings, addressing both the practical applications and the limitations of the study. Finally, @sec-conclusion concludes the paper and suggests directions for future research.

@@

## Overview
Over the ten-year period, the dataset reveals the nature and frequency of three primary types of crisis events: Overdose, Person in Crisis, and Suicide-related, as summarized in @tbl-summary. Among these, Person in Crisis events are by far the most frequent, comprising more than half of the total incidents.

@@

In terms of location, areas like Moss Park and Downtown Yonge East consistently report high numbers of crisis calls, particularly for overdoses. These findings suggest the need for focused interventions in these neighborhoods, potentially influenced by socioeconomic conditions or population density. Recognizing both the temporal and geographic patterns of crisis incidents is key to enhancing emergency response strategies and supporting the areas most impacted.

@@

The dataset also includes additional information that could deepen the analysis, such as police division details, whether an official occurrence was logged, and whether a Mental Health Act (MHA) apprehension occurred. Future studies could examine how these elements influence crisis resolution. Moreover, although this dataset focuses on three crisis types, incorporating a broader range of emergency call categories could provide a more detailed understanding of Toronto's crisis response landscape.

# Acknowledgements

@@

# Appendix

>>>>

The main sections are organized as follows: 
@sec-data provides a detailed description of the data and its sources, including the context and ethical considerations. @sec-results presents the results using visualizations as well as discusses the implications of the findings, addressing both the practical applications and the limitations of the study. @sec-conclusion concludes the paper and suggests directions for future research.

@@

## Result Overview
Over the ten-year period, the dataset reveals the nature and frequency of three primary types of crisis events: Overdose, Person in Crisis, and Suicide-related, as summarized in @tbl-summary. Among these, Person in Crisis events are by far the most frequent, comprising more than half of the total incidents.

@@

In terms of location, areas like Moss Park and Downtown Yonge East consistently report high numbers of crisis calls, particularly for overdoses. These findings suggest the need for focused interventions in these neighborhoods, potentially influenced by socioeconomic conditions or population density. Notably, Moss Park is one of Toronto’s main encampment sites, and the research on pandemic encampments by @edhi2022pandemic provides a complementary perspective on the broader social context within which these crises occur. The study highlights that the City of Toronto has been experiencing one of its deepest housing crises, with homeless encampments serving as sites of resistance against systemic violence and insufficient interventions. The displacement and criminalization of unhoused residents during the pandemic underscore the intersectionality of housing insecurity and mental health crises. As such, addressing the root causes of homelessness, including the lack of affordable housing and supportive services, could contribute to reducing the frequency of crisis calls, especially in areas with high concentrations of vulnerable populations. Integrating a human rights-based approach to housing and crisis management, as advocated by Edhi, can provide a more holistic framework for responding to and preventing these emergencies, thereby fostering a more equitable and resilient urban environment.

@@

In terms of future work, the dataset also includes additional information that could deepen the analysis, such as police division details, whether an official occurrence was logged, and whether a Mental Health Act (MHA) apprehension occurred. Future studies could examine how these elements influence crisis resolution. Moreover, although this dataset focuses on three crisis types, incorporating a broader range of emergency call categories could provide a more detailed understanding of Toronto's crisis response landscape.

# Appendix {#sec-app}

## Acknowledgements

@@
